\section*{Nomenclature}

\subsection*{Constants of nature}
\begin{itemize}
    \item[$m_e$] - Mass of electron.\\
        $m_e = 9.10938356\cdot10^{-31}\unit{kg}$.
    \item[$m_\mathrm{H}$] - Mass of hydrogen atom. \\
        $m_\mathrm{H} = 1.6735575\cdot10^{-27}\unit{kg}$.
    \item[$G$] - Gravitational constant.\\
            $G=6.67430$\stdf{-11}$\unit{m}^3\unit{kg}^{-1}\unit{s}^{-2}.$
    \item[$k_B$] - Boltzmann constant.\\
            $k_B = 1.38064852\cdot10^{-23}\unit{m}^2\unit{kg}\unit{s}^{-2}\unit{K}^{-1}.$
    \item[$\hbar$] - Reduced Planck constant.\\
            $\hbar = 1.054571817\cdot10^{-34}\unit{J}\unit{s}^{-1}$.
    \item[$c$] - Speed of light in vacuum.\\
            $c=2.99792458 \cdot10^8\unit{m}\unit{s}^{-1}$.
    \item[$\sigma_T$] - Thomson cross section.\\
            $\sigma_T = 6.6524587158\cdot10^{-29} \unit{m}^2$.
    \item[$\alpha$] - Fine structure constant.\\
        $\alpha = \frac{m_ec}{\hbar}\sqrt{\frac{3\sigma_T}{8\pi}}$
\end{itemize}

\subsection*{Cosmological parameters}
\begin{itemize}
    \item[$G_\munu$] - Einstein tensor.
    \item[$T_\munu$] - Stress-energy tensor.
    \item[$H$] - Hubble parameter.
%     \item[$H_0$] - Hubble constant \checkthis{fill in stuff}.
    \item[$\Hp$]  - Conformal Hubble parameter.
    \item[$T_{\mathrm{CMB}0}$] - Temperature of CMB today.
    \item[$a$] - Scale factor.
    \item[$x$] - Logarithm of scale factor.
    \item[$t$] - Cosmic time.
    \item[$z$] - Redshift.
    \item[$\eta$] - Conformal time.
    \item[$\chi$] - Co-moving distance. 
    \item[p] - Pressure.
    \item[$\rho$] - Density.
    \item[$r$] - Radial distance.
    \item[$d_A$] - Angular diameter distance.
    \item[$d_L$] - Luminosity distance.
    \item[] --------------------------
    \item[$n_e$] - Electron density.
    \item[$n_b$] - Baryon density. 
    \item[$X_e$] - Free electron fraction.
    \item[$\tau$] - Optical depth.
    \item[$\tilde{g}$] - Visibility function. 
    \item[$s$] - Sound horizon.
    \item[$r_s$] - Sound horizon at decoupling. 
    \item[$c_s$] - Wave propagation speed.
    \item[] --------------------------
    \item[$\Psi$] - Newtonian potential perturbation to the metric. 
    \item[$\Phi$] - Spatial curvature perturbation to the metric.
    \item[$P^\mu$]- Comoving momentum four vector.
    \item[$\mathcal{P}_l$] - Legendre polynomial of order $l$.
    \item[$k$] - Fourier mode $k=\vec{k}$ of wave vector $\vec{k}$. 
    \item[$\T$] - Photon perturbation.
    \item[$\T_l$] - $l$-th order multipole expansion term of $\T$.
    \item[$v_b$] - Baryon bulk velocity.
    \item[$v_c$] - Cold dark matter bulk velocity.
    \item[$v_\gamma$] - Photon bulk velocity. 
    \item[$\delta_b$] - Baryon overdensity.
    \item[$\delta_c$] - Cold dark matter overdensity.         
    \item[$\delta_\gamma$] - Photon overdensity.
    \item[$\phi$] - Inflaton field. 
    \item[$\rho_\phi$] - Inflaton field density.
    \item[$p_\phi$] - Inflaton field pressure.
    \item[$\epsilon_\mathrm{sr}, \delta_\mathrm{sr}$] - Slow roll parameters.  
    \item[$\mathcal{R}$] - Curvature perturbation.
    \item[$\mathcal{S}$] - Source function.  
    \item[] --------------------------
    \item[$\To$] - Observed temperature anisotropies.  
    \item[$Y_{lm}$] - Spherical harmonic functions.  
    \item[$a_{lm}$] - Coefficients in spherical harmonic expansion. 
    \item[$C_l$] - Angular power spectrum.
    \item[$P_\mathrm{prim}$] - Primordial power spectrum.
    \item[$n_s$] - Spectral index.
    \item[$A_s$] - Primordial amplitude.  
    \item[$\Delta_\mathrm{M}$] - Total matter contrast.
    \item[$P$] - Matter power spectrum. 
    \item[$j_l$] - Spherical Bessel function.  
\end{itemize}

\subsection*{Density parameters}
Density parameter $\O_X = \rho_X/\rho_\mathrm{crit}$ where $\rho_X$ is the density and $\rho_c=8\pi G/3H^2$ the critical density. $X$ can take the following values:
\begin{itemize}
    \item[$b$] - Baryons.
    \item[$c$] - Cold dark matter.
    \item[$\gamma$] - Electromagnetic radiation.
    \item[$\nu$] - Neutrinos.
    \item[$k$] - Spatial curvature.
    \item[$\Lambda$] - Cosmological constant.
\end{itemize}
A $0$ in the subscript indicates the present day value. 

\subsection*{Fiducial cosmology}
The fiducial cosmology used throughout this project is based on the observational data obtained by ~\cite{Planck2020}:
\begin{equation*}
        \begin{split}
                h &= 0.67, \\
                T_\mathrm{CMB0} &= 2.7255\unit{K}, \\
                N_\mathrm{eff} &= 3.046, \\
                \O_\mathrm{b0} &= 0.05, \\
                \O_\mathrm{c0} &= 0.267, \\
                \O_{k0} &= 0, \\
                \O_{\nu0} &= N_\mathrm{eff} \cdot \frac{7}{8}\left(\frac{4}{11}\right)^{4/3}\O_{\gamma0}, \\
                \O_{\Lambda0} &= 1- (\O_{k0} + \O_\mathrm{b0} + \O_\mathrm{CDM0} + \O_{\gamma0} + \O_{\nu0}),\\
                \O_\mathrm{M0} &= \O_{b0} + \O_\mathrm{CDM0}, \\
                \O_\mathrm{rad} &= \O_{\gamma0} + \O_{\nu0},\\
                n_s &= 0.965, \\
                A_s &= 2.1\cdot10^{-9}.
        \end{split}
\end{equation*}