\section{Recombination History}\label{sec:m2}

The main goal of this section is to investigate the recombination history of the universe. This can be explained as the point in time when photons decouple from the equilibrium of the opaque early universe.  This is known as the \textit{time of last scattering},\footnote{Which is exactly what the name suggests.} and these photons are what we today observe as the CMB. This period of the history of the universe is thus crucial for understanding the CMB. 

We will start by calculating the free \textit{electron fraction} $X_e$, from which we may find the \textit{optical depth} $\tau$. This again enables us to compute the \textit{visibility function}, $g$, and the \textit{sound horizon}, $s$. The latter will be of great importance later. 

Recombination happens because the expansion of the Universe cools it down, making the photons less energetic, which in turn make each interaction in the primordial plasma less energetic. At some point, hydrogen atoms are able to form, reducing the number of free electron, hence reducing photon interactions, until they scatter for the last time. We will determine the time of recombination from the free electron fraction, which indirectly tell us how large portion of the free electrons have (re)-combined.\footnote{As with any (hopefully good) article on the subject, we ought to say that recombination is a funny wording, as this is the first time in the history of the Universe that protons and electrons combine to form hydrogen.} Due to the decrease of free electrons, photons interact less with them. At some point, photons scatter for the last time, and this information is encapsulated in the visibility function. 


\input{sections/M2/m2_theory}
\input{sections/M2/m2_methods}
\input{sections/M2/m2_results}

